Concurrent Programming in Harmony

### 2021-08-07 book/paper.tex
Working on chain replication chapter

@@

@@

@@

>>>>

\newglossaryentry{replication}
  name=replication,
  description={maintains multiple copies of some resource to improve availability in the face of failures}

@@

\newglossaryentry{state machine replication}
  name=state machine replication,
  description={is a replication technique in which a collection of deterministic state machines process the same inputs in the same order}

@@

\item \autoref{ch:chain} describes \emph{state machine replication} and
the \emph{chain replication} protocol to support replication.

@@

\chapter{Chain Replication}
\label{ch:chain}

\begin{figure}
\begin{code}
\harmonysource{chain}
\end{code}
\caption{\harmonylink{code/chain.hny} Chain Replication.}
\label{fig:chain}
\end{figure}

\glsadd{replication}%
\index{replication}%
As you have probably experienced, computers can crash.
If you are running a web service, you may not be able to afford
a long outage.  If you are running software that flies a plane,
then an outage for any length of time could lead to a disaster.
To deal with service outages caused by computers crashing, you may want
to \emph{replicate} the service onto multiple computers.  As long as
one of the computers survives, the service remains available.

\glsadd{state machine replication}%
\index{state machine replication}%
Besides availability, it is usually important that the replicated
services acts as if it were a single one.  This requires that the
replicas of the service must coordinate their actions.
One general approach is the
\emph{Replicated State Machine Approach}~\cite{Lam78,S90}.
In this approach, you model your service as a deterministic state
machine.  The replicas each run a copy of the state machine, started
in the same state.  As long as the replicas handle the same inputs
in the same order, determinism guarantees that they produce the same
outputs in the same order.
The trick then is to ensure that all replicas handle the
same requests in the same order and to do so in a way that continues
to work even if some subset of replicas crash.

\emph{Chain Replication}~\cite{vRS04} is a protocol to do just this.
In Chain Replication, the replicas are organized in a linear chain.
Each replica monitors its direct predecessor in the chain.  If it has
no predecessors, we call it the \emph{head}.  The head may change over
time, as a head may crash and replaced by another.
We model the state machine as a \emph{history}: a sequence of operations.
Only the head is allowed to introduce new updates.
When it does so, it advertises the new update along with the position
in the history.  When the direct successor receives the next update
it is expecting, it appends the update to its history and likely
advertises the new update.  When the update is applied to the history
of the \emph{tail} replica (the replica that has no live successors),
then the update is considered final.

So, when a replica fails, its successors should find out about it.
In practice, one server can detect the failure of another server by
pinging it.  If a server does not receive a response to its ping within
some maximum amount of time, then the server considers its peer crashed.
Note that this, in general is not a safe thing to do---the network or
the peer may be temporarily slow but the peer is not necessarily crashed
when the timer expires.
Nonetheless, we will assume here that failure detection does not make
mistakes and that eventually every failure is detected.
This is sometimes called the \emph{Fail-Stop} failure model~\cite{SS83},
which is distinct from the often more realistic \emph{Crash} failure model
where processes can crash but accurate detection is not available.

\autoref{fig:chain} shows an implemenation of chain replication.
